---
title: "Exploratory Analysis"
author: "Nikolas Krstic"
date: "September 15, 2017"
output: github_document
keep_md: yes
---

## Load necessary packages/data

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
```


## Explore full "gapminder" dataset


__Gapminder is a list__
```{r}
typeof(gapminder)
```

__Gapminder is a tibble (can also be considered a data.frame)__
```{r}
class(gapminder)
```

__Gapminder has 6 columns/variables__
```{r}
ncol(gapminder)
```

__Gapminder has 1704 rows/observations__
```{r}
nrow(gapminder)
```

__Another few functions that can be applied to get the gapminder dimensions. Particularly useful if handling data that's not necessarily two dimensional (arrays, etc.). Also, "str" provides information on the class of each variable.__
```{r}
dim(gapminder)
str(gapminder)
```

__The data type of each variable:__

* country: Factor
* continent: Factor
* year: Integer
* lifeExp: Double
* pop: Integer
* gdpPercap: Double


## Explore individual variables

__Frequency table of "continent" variable__
```{r}
kable(t(as.matrix(table(gapminder$continent))))
```


__Frequency table of "country" variable__
```{r}
kable(t(as.matrix(table(gapminder$country))))
```

__Frequency table of "year" variable__
```{r}
kable(t(as.matrix(table(gapminder$year))))
```

__Summary Statistics of "lifeExp" variable__
```{r}
LE_SS = gapminder %>%
  summarize(Minimum = min(lifeExp), Maximum = max(lifeExp),
            Mean = mean(lifeExp), SD = sd(lifeExp))

kable(LE_SS)
```


__Summary Statistics of "gdpPercap" variable__
```{r}
GPC_SS = gapminder %>%
  summarize(Minimum = min(gdpPercap), Maximum = max(gdpPercap),
            Mean = mean(gdpPercap), SD = sd(gdpPercap))

kable(GPC_SS)
```


__Summary Statistics of "pop" variable__
```{r}
POP_SS = gapminder %>%
  summarize(Minimum = min(pop), Maximum = max(pop),
            Mean = mean(pop), SD = sd(pop))

kable(POP_SS)
```



