---
title: "Exploratory Analysis"
author: "Nikolas Krstic"
date: "September 15, 2017"
output: github_document
keep_md: yes
---

## Load necessary packages/data

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
```


## Explore full "gapminder" dataset


#### Gapminder is a list
```{r}
typeof(gapminder)
```

#### Gapminder is a tibble (can also be considered a data.frame)
```{r}
class(gapminder)
```

#### Gapminder has 6 columns/variables
```{r}
ncol(gapminder)
```

#### Gapminder has 1704 rows/observations
```{r}
nrow(gapminder)
```

__Another few functions that can be applied to get the gapminder dimensions. Particularly useful if handling data that's not necessarily two dimensional (arrays, etc.). Also, "str" provides information on the class of each variable.__
```{r}
dim(gapminder)
str(gapminder)
```

#### The data type of each variable:

* country: Factor
* continent: Factor
* year: Integer
* lifeExp: Double
* pop: Integer
* gdpPercap: Double


## Explore individual variables

#### Frequency table of "continent" variable
```{r}
kable(t(as.matrix(table(gapminder$continent))))
```


#### Frequency table of "country" variable
```{r}
kable(t(as.matrix(table(gapminder$country))))
```

#### Frequency table of "year" variable
```{r}
kable(t(as.matrix(table(gapminder$year))))
```

#### Summary Statistics of "lifeExp" variable
```{r}
LE_SS = gapminder %>%
  summarize(Minimum = min(lifeExp), Maximum = max(lifeExp),
            Mean = mean(lifeExp), SD = sd(lifeExp))

kable(LE_SS)
```


#### Summary Statistics of "gdpPercap" variable
```{r}
GPC_SS = gapminder %>%
  summarize(Minimum = min(gdpPercap), Maximum = max(gdpPercap),
            Mean = mean(gdpPercap), SD = sd(gdpPercap))

kable(GPC_SS)
```


#### Summary Statistics of "pop" variable
```{r}
POP_SS = gapminder %>%
  summarize(Minimum = min(pop), Maximum = max(pop),
            Mean = mean(pop), SD = sd(pop))

kable(POP_SS)
```



