---
title: "Singer_Locations_Factors_and_Writing"
author: "Nikolas Krstic"
date: "October 13, 2017"
output: github_document
keep_md: yes
---

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(singer))
suppressPackageStartupMessages(library(forcats))
```


## Factor Management

```{r}
data("singer_locations")
```

We're going to factorise on the following variables: artist_name, name and city. This is because of the following information we have on the unique values of each of the categorical variables (that are not identifiers):
```{r}
#10100
nrow(singer_locations)
#9799, too high and close to nrow value, not useful to factorise
length(unique(singer_locations$title))
#9049, too high again
length(unique(singer_locations$release))
#7498, acceptable because of absence of NAs
length(unique(singer_locations$artist_name))
#2913, good enough
length(unique(singer_locations$name))
#1317, good enough as well
length(unique(singer_locations$release))
```

#### Check for NA count
```{r}
#5968
sum(is.na(singer_locations$name))
#0
sum(is.na(singer_locations$artist_name))
#5971
sum(is.na(singer_locations$city))
```

### Factorisation
```{r}
#Replace any NAs present with "Missing Information", to allow us to use as_factor
sl_new <- singer_locations %>%
  mutate(name = ifelse(is.na(singer_locations$name), "Missing Information", singer_locations$name), artist_name = ifelse(is.na(singer_locations$artist_name), "Missing Information", singer_locations$artist_name), city = ifelse(is.na(singer_locations$city), "Missing Information", singer_locations$city))

sl_new <- sl_new %>%
  mutate(name_factor = as_factor(sl_new$name), artist_name_factor = as_factor(sl_new$artist_name), city_factor = as_factor(sl_new$city))
```

It should be noted that as_factor sets the levels in the order that they first appear in the data, while as.factor sets the levels in alphabetical order. A serious drawback to as_factor though is that the NAs within a variable hinder its ability to factorise. This is not a problem with as.factor, and usually you will end up changing the order of factor levels, regardless of whether you use either function.

#### Name of specific singer
```{r}
#Number of levels
nlevels(sl_new$name_factor)

#Some sample levels
head(levels(sl_new$name_factor), n=10)
```


#### Artist name (can be band/singer/artist/etc.)
```{r}
nlevels(sl_new$artist_name_factor)

head(levels(sl_new$artist_name_factor), n=10)
```

#### City
```{r}
nlevels(sl_new$city_factor)

head(levels(sl_new$city_factor), n=10)
```


### Removal of year 0 from dataset and drop factor levels
```{r}
sl_new2 <- sl_new %>%
  filter(year != 0) %>%
  droplevels()

nrow(sl_new2)
```

We had 100 observations that were year 0. We examine how the factors have changed as well.

#### Name of specific singer
```{r}
#Number of levels
nlevels(sl_new2$name_factor)

#Some sample levels
head(levels(sl_new2$name_factor), n=10)
```

We have 34 levels that were dropped.

#### Artist name (can be band/singer/artist/etc.)
```{r}
nlevels(sl_new2$artist_name_factor)

head(levels(sl_new2$artist_name_factor), n=10)
```

We have 90 levels that were dropped.

#### City
```{r}
nlevels(sl_new2$city_factor)

head(levels(sl_new2$city_factor), n=10)
```

We have 8 levels that were dropped.

### Reordering the levels of arist_name_factor by maximum artist_familiarity

```{r}
sl_new3 <- sl_new2 %>%
  mutate(artist_name_factor = fct_reorder(artist_name_factor, artist_familiarity, max))

head(levels(sl_new3$artist_name_factor), n=10)
```
Factor levels have been reordered.

### Writing and Reading to File

```{r}
sl_new4 <- sl_new3 %>%
  mutate(city_factor = fct_reorder(city_factor, artist_familiarity, max)) %>%
  group_by(city_factor) %>%
  summarize(mean_artist_hotttnesss = mean(artist_hotttnesss))

head(levels(sl_new4$city_factor), n=10)

write_csv(sl_new4, paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))
sl_new5 <- read_csv(paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))

str(sl_new5)
```

It appears that the factor was not preserved with write_csv

```{r}
saveRDS(sl_new4, paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))
sl_new5 <- readRDS(paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))

str(sl_new5)
head(levels(sl_new5$city_factor), n=10)
```
It appears that the factor (and its reordered levels) were preserved with saveRDS

```{r}
dput(sl_new4, paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))
sl_new5 <- dget(paste(getwd(), "/Data/Artist_Hotness_Data.csv", sep=""))

str(sl_new5)
head(levels(sl_new5$city_factor), n=10)
```
It appears that the factor (and its reordered levels) were preserved with dput

### Visualization Design

```{r}
sl_new6 <- sl_new3 %>%
  group_by(city_factor) %>%
  summarize(mean_artist_familiarity = mean(artist_familiarity)) %>%
  mutate(city_factor = fct_reorder(city_factor, mean_artist_familiarity, max)) %>%
  filter(mean_artist_familiarity>0.85)

sl_new6 %>%
  ggplot(aes(x=city_factor, y=mean_artist_familiarity)) +
  theme_bw() +
  geom_point() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(x="City", y="Artist Hotness", title="Cities with the Most Familiar Artists")
```

The factor along the x-axis appears to be ordered in the figure appropriately based on our reordering (based on maximum of mean artist familiarity).

```{r}
sl_new7 <- sl_new3 %>%
  group_by(city_factor) %>%
  summarize(mean_artist_familiarity = mean(artist_familiarity)) %>%
  arrange(mean_artist_familiarity) %>%
  filter(mean_artist_familiarity>0.85)

sl_new7 %>%
  ggplot(aes(x=city_factor, y=mean_artist_familiarity)) +
  theme_bw() +
  geom_point() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(x="City", y="Artist Hotness", title="Cities with the Most Familiar Artists")
```

However, in contrast, using only "arrange" to attempt to reorganize the appearance of the figure doesn't work

```{r}
#Use the previous dataframe that had the reordered city factor
new_plot = sl_new6 %>%
  arrange(mean_artist_familiarity) %>%
  ggplot(aes(x=city_factor, y=mean_artist_familiarity)) +
  theme_bw() +
  geom_point() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(x="City", y="Artist Hotness", title="Cities with the Most Familiar Artists")

new_plot
```

Combining "arrange" with reordering of the factor does result in the same plot as just performing the reordering, so we can conclude that arranging the data has no effect on the plot.

### Writing Figures to File

```{r}
ggsave(paste(getwd(), "/Plots/Mean_Artist_Familiarity_Plot.pdf", sep=""), plot=new_plot, height=6)
```


The embed doesn't seem to work, and I'm not entirely sure why. I would rather not put my local drive information in order to get this to work, however. You can find the plot in the corresponding folder, nevertheless

Saving our plot in the vector format (.pdf, .eps) because it provides the best resolution for graphs in general. Raster format is useful for photographs and other detailed visuals, but can potentially provide poorer resolution.

Went with most of the default options because they are often the best.

By not specifying the plot argument, ggsave will automatically save the last plot that was displayed, and therefore it's important to specify the plot you would like to save.

### But I want to do More:

```{r}
sl_new8 <- sl_new3 %>%
  filter(city_factor %in% c(levels(city_factor)[1:10])) %>%
  droplevels() %>%
  mutate(state = fct_recode(city_factor, "NA" = "Missing Information", "Illinois" = "Chicago, IL", "New York" = "New York, NY", "Michigan" = "Detroit, MI", "California" = "Oxnard, CA", "California" = "Los Angeles, CA", "New York" = "Staten Island, NY"))

kable(head(sl_new8, n=30))
```

Created the factor for the states of a handful of cities. May become tedious to have to recode every single factor level when generating the new factor.


### Report on Process

Encountered a few challenges, namely trying to embed an image within the report. I was unsuccessful and I'm not sure if it's required that I point to the file directly on my local drive (that doesn't make sense to me). However, it insists the file doesn't exist if I don't specify the full path (even with the work directory set to within this project)

Otherwise, I feel like I had a good grasp on how factors worked before this week, but now I have a better understanding of how to manipulate them with dplyr, which I was originally inexperienced with before this class.

Also, it seems that write_csv is generally better than write.csv, which I used fairly often. What's good about it is that it doesn't write the row names as well. This is useful because consistent reading and writing without proper data cleaning would result in multiple columns of row names, which was annoying.





